Solar-wind predictions for the Parker Solar Probe orbit

### 2017-11-17 paperMVVB_language_editor/31831_corr.tex
implemented the *.tex file from the language editor

@@

\begin{document}
        \title{Solar-wind predictions for the Parker Solar Probe orbit}
        \subtitle{Near-Sun extrapolations derived from an empirical solar-wind model based on Helios and OMNI observations}
        \author{M.~S.~Venzmer
        \and V.~Bothmer}
        \institute{University of Goettingen, Institute for Astrophysics, Friedrich-Hund-Platz~1, 37077~Göttingen, Germany}
        \date{Received 25 August 2017; accepted date}

@@

        The study presented in this paper is undertaken in the Coronagraphic German And US SolarProbePlus Survey (CGAUSS) project, which is the German contribution to the PSP mission as part of the Wide-field Imager for Solar PRobe (WISPR). WISPR will contribute to the PSP science goals by deriving the three-dimensional structure of the solar corona through which the in situ measurements are made to determine the sources of the solar wind. It will provide density power spectra over a wide range of structures (e.g., streamers, pseudostreamers and equatorial coronal holes) for determining the roles of turbulence, waves, and pressure-balanced structures in the solar wind. It will also measure the physical properties, such as speed and density jumps of SEP-producing shocks and their coronal mass ejection (CME) drivers as they evolve in the corona and inner heliosphere \citep{Vourlidas2016}.
        In order to help optimize the WISPR and PSP preplanning of the science operations, knowledge of the expected solar-wind environment is needed. For this purpose the solar-wind environment is extrapolated down to the closest perihelion of \SI{9.86}{\Rs} distance to the Sun using in situ solar-wind data from the Helios probes and near \SI{1}{\au} data from various satellites compiled in the OMNI solar-wind database.

@@

        The OMNI~2 data set \citep{King2005} combines solar-wind magnetic field and plasma data collected by various satellites since 1963, currently by WIND and by ACE. This intercalibrated multi-spacecraft data is time-shifted to the nose of the Earth’s bow shock. The data is obtained from the OMNIWeb interface\footnote{\url{http://omniweb.gsfc.nasa.gov/}} at NASA's Space Physics Data Facility (SPDF), Goddard Space Flight Center (GSFC).
        In this study the whole hourly data until 31 December 2016 is used, starting from 27 November 1963 (for the temperature from 26 July 1965). The data coverage of the different parameters is in the range \SIrange{67}{74}{\percent},  corresponding to a total duration of 36--40~years.
        We note that a test-comparison of hourly averaged data with higher-time-resolution data for the available shorter time span 1981--2016 did not show significant differences in our results.
        According to the OMNI data precision and maximal parameter ranges we specify bin sizes of \SI{0.5}{\nT} for the magnetic field strength, \SI{10}{\km\per\s} for the velocity, \SI{1}{\per\cm\cubed} for the density and \SI{10000}{\K} for the temperature. The frequency distributions of the solar-wind magnetic field strength, proton velocity, density and temperature are shown in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}.
        The solar-wind magnetic field strength is in the range \SIrange{0.4}{62}{nT}, the velocity in the range \SIrange{156}{1189}{\km\per\s}, the density in the range \SIrange{0}{117}{\per\cm\cubed}, the temperature in the range \SIrange{3450}{6.63e6}{\K}, and the mean data values are at \SI{6.28}{\nT}, \SI{436}{\km\per\s}, \SI{6.8}{\per\cm\cubed} and \SI{1.05e5}{\K}. These ranges and mean values are as statistically expected from previous analyses of near \SI{1}{\au} solar-wind data (e.g., Table~3.3 in \citet[p.~39]{Bothmer2007}).
        Much higher or lower peak values at \SI{1}{\au} have been observed in extraordinary events, such as the 23~July 2012 CME with a speed of over \SI{2000}{\km\per\s} and a peak field strength of about \SI{100}{\nT} that was observed by STEREO~A \citep{Russell2013}, or the solar-wind disappearance event observed in May 1999 with density values even down to \SI{0.2}{\per\cm\cubed} \citep{Lazarus2000}.

@@

        Again, the solar-wind velocity needs a special treatment because of the application of the double lognormal distribution (\ref{eq:double_lognormal_fit_function}). Since it is well known that slow and fast solar-wind stream occurrence rates follow the solar cycle, we keep the two velocity components' positions SSN-independent ($x_\text{med} =  b_\text{med}$) and vary instead their balance with the SSN:
        \begin{align}
                c(ssn) &= c_a \cdot ssn + c_b\,.        \label{eq:balance_with_ssn}
        \end{align}
        The fit result (see Table~\ref{tab:ssn_fit_parameters}) yields a model in which three years after solar cycle minimum (SSN of zero) the contribution of slow solar wind to the overall solar wind distribution reaches a maximum value (about \SI{64}{\percent}) and decreases with increasing SSN as shown in Fig.~\ref{fig:Vdbl_SSN_ratio_f_plot}.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{figures/Vdbl_SSN_ratio_f_plot.pdf}}
                \caption{Ratio of slow to fast solar wind for a SSN lagged by 3~years. The yearly ratios (+) and their weighted linear fit (solid line) are obtained from OMNI data with a threshold velocity of $v_\text{th} = \SI{400}{\km\per\s}$. The error bars denote the SSN standard deviation and the relative weight from the yearly data coverage. The model's balance parameter (\ref{eq:balance_with_ssn}) and derived ratio (same threshold) are plotted as dashed and dotted lines.}
                \label{fig:Vdbl_SSN_ratio_f_plot}
        \end{figure}

@@

                \caption{Fit coefficients for the distance-dependent single lognormal function, based on Equation (\ref{eq:single_lognormal_fit_function}) combined with (\ref{eq:power_function}) from the combined Helios data. Regarding the velocity, the double lognormal function (\ref{eq:double_lognormal_fit_function}) is used instead. The numbers in parentheses are the errors on the corresponding last digits of the quoted value. They are calculated from the estimated standard deviations of the fit parameters. The seasonal variations are calculated from Earth's orbital solar distance variation and the derived exponents.}

@@

                \includegraphics[width=18cm]{figures/sw_extrapolation_ssn_f_plot.pdf}
                \caption{Radial extrapolation of the solar-wind parameters to the PSP orbit region. The median values from the models, obtained from Helios and OMNI measurements, are extrapolated to the PSP region for SSN values between solar minimum and maximum, that is, $0 \le ssn \le 200$. The lower edges of the shaded areas correspond to solar minimum, the upper edges to solar maximum. Also plotted are the radial dependencies of the analytic DQCS magnetic field model for solar minimum from \citet{Banaszkiewicz1998}, the slow wind velocity models from \citet{Sheeley1997} and \citet{Wang2000}, the density model from \citet{Leblanc1998} and the range of temperature measurements from \citet{Billings1959} and \citet{Liebenberg1975}.}

@@

        For determining solar-activity- and solar-distance-dependent relations for the median and mean solar-wind values, we could have used the simpler approach of combining the radial dependence of averaged Helios data with averaged \SI{1}{\au} OMNI data scaled with the SSN. It is expected that the results of a simpler analysis would have similar distance scaling results, as can be inferred from the exponents in Tables~\ref{tab:mean_median_fit_parameter} and \ref{tab:extrapolation_model_fit_parameters}. However, in our study we are not only interested in averages but rather in bulk distributions, that is, the whole range of values that might occur. For the determination of the frequency distributions the use of the more complex fit model is important, because the distance between median and mean values determines the width of the lognormal distributions.

>>>>

%

@@

\begin{document}

        \title{Solar-wind predictions for the Parker Solar Probe orbit}
        \subtitle{Near-Sun extrapolations derived from an empirical solar-wind model based on Helios and OMNI observations}
        \author{M.~S.~Venzmer
        \and V.~Bothmer}
        \institute{University of Goettingen, Institute for Astrophysics, Friedrich-Hund-Platz~1, 37077~Göttingen, Germany}
        \date{Received 25 August 2017; accepted date}

@@

        The study presented in this paper is undertaken in the Coronagraphic German And US SolarProbePlus Survey (CGAUSS) project, which is the German contribution to the PSP mission as part of the Wide-field Imager for Solar PRobe (WISPR). WISPR will contribute to the PSP science goals by deriving the three-dimaensional (3D) structure of the solar corona through which the in situ measurements are made to determine the sources of the solar wind. It will provide density power spectra over a wide range of structures (e.g., streamers, pseudostreamers and equatorial coronal holes) for determining the roles of turbulence, waves, and pressure-balanced structures in the solar wind. It will also measure the physical properties, such as speed and density jumps of SEP-producing shocks and their coronal mass ejection (CME) drivers as they evolve in the corona and inner heliosphere \citep{Vourlidas2016}.
        In order to help optimize the WISPR and PSP preplanning of the science operations, knowledge of the expected solar-wind environment is needed. For this purpose the solar-wind environment is extrapolated down to the closest perihelion of \SI{9.86}{\Rs} distance to the Sun using in situ solar-wind data from the Helios probes and near \SI{1}{\au} data from various satellites compiled in the OMNI solar-wind database.

@@

        The OMNI~2 data set \citep{King2005} combines solar-wind magnetic field and plasma data collected by various satellites since 1963, currently by WIND and by ACE. This intercalibrated multi-spacecraft data is time-shifted to the nose of the Earth’s bow shock. The data is obtained from the OMNIWeb interface\footnote{\url{http://omniweb.gsfc.nasa.gov/}} at NASA's Space Physics Data Facility (SPDF), Goddard Space Flight Center (GSFC).
        In this study the whole hourly data until 31 December 2016 is used, starting from 27 November 1963 (for the temperature from 26 July 1965). The data coverage of the different parameters is in the range \SIrange{67}{74}{\percent},  corresponding to a total duration of 36--40~years.
        We note that a test-comparison of hourly averaged data points with higher-time-resolution data for the available shorter time span 1981--2016 did not show significant differences in our results.        According to the OMNI data precision and maximal parameter ranges we specify bin sizes of \SI{0.5}{\nT} for the magnetic field strength, \SI{10}{\km\per\s} for the velocity, \SI{1}{\per\cm\cubed} for the density and \SI{10000}{\K} for the temperature. The frequency distributions of the solar-wind magnetic field strength, proton velocity, density and temperature are shown in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}.
        The solar-wind magnetic field strength is in the range \SIrange{0.4}{62}{nT}, the velocity in the range \SIrange{156}{1189}{\km\per\s}, the density in the range \SIrange{0}{117}{\per\cm\cubed}, the temperature in the range \SIrange{3450}{6.63e6}{\K}, and the mean data values are at \SI{6.28}{\nT}, \SI{436}{\km\per\s}, \SI{6.8}{\per\cm\cubed} and \SI{1.05e5}{\K}. These ranges and mean values are as statistically expected from previous analyses of near \SI{1}{\au} solar-wind data (e.g., Table~3.3 in \citet[p.~39]{Bothmer2007}).
        Much higher or lower peak values at \SI{1}{\au} have been observed in extraordinary events, such as the 23~July 2012 CME with a speed of over \SI{2000}{\km\per\s} and a peak field strength of about \SI{100}{\nT} that was observed by STEREO~A \citep{Russell2013}, or the solar-wind disappearance event observed in May 1999 with density values even down to \SI{0.2}{\per\cm\cubed} \citep{Lazarus2000}.

@@

        Again, the solar-wind velocity needs a special treatment because of the application of the double lognormal distribution (\ref{eq:double_lognormal_fit_function}). Since it is well known that slow and fast solar-wind stream occurrence rates follow the solar cycle, we keep the two velocity components' positions SSN-independent ($x_\text{med} =  b_\text{med}$) and vary instead their balance with the SSN:
        \begin{align}
                c(ssn) &= c_a \cdot ssn + c_b\,.        \label{eq:balance_with_ssn}
        \end{align}
        The fit result (see Table~\ref{tab:ssn_fit_parameters}) yields a model in which three years after solar cycle minimum (SSN of zero) the contribution of slow solar wind to the overall solar wind distribution reaches a maximum value (about \SI{64}{\percent}) and decreases with increasing SSN as shown in Fig.~\ref{fig:Vdbl_SSN_ratio_f_plot}.
        \begin{figure}
                \resizebox{\hsize}{!}{\includegraphics{figures/Vdbl_SSN_ratio_f_plot.pdf}}
                \caption{Ratio of slow to fast solar wind for a\LEt{Please insert missing text.} by 3~years lagged SSN. The yearly ratios (+) and their weighted linear fit (solid line) are obtained from OMNI data with a threshold velocity of $v_\text{th} = \SI{400}{\km\per\s}$. The error bars denote the SSN standard deviation and the relative weight from the yearly data coverage. The model's balance parameter (\ref{eq:balance_with_ssn}) and derived ratio (same threshold) are plotted as dashed and dotted lines.}
                \label{fig:Vdbl_SSN_ratio_f_plot}
        \end{figure}

@@

                \caption{Fit coefficients for the distance-dependent single lognormal function, based on Equation (\ref{eq:single_lognormal_fit_function}) combined with (\ref{eq:power_function}), respectively double lognormal function (\ref{eq:double_lognormal_fit_function}) for the velocity from the combined Helios data\LEt{The meaning of this second part of the sentence is unclear; please rephrase for improved clarity.}. The numbers in parentheses are the errors on the corresponding last digits of the quoted value. They are calculated from the estimated standard deviations of the fit parameters. The seasonal variations are calculated from Earth's orbital solar distance variation and the derived exponents.}

@@

                \includegraphics[width=18cm]{figures/sw_extrapolation_ssn_f_plot.pdf}           \caption{Radial extrapolation of the solar-wind parameters to the PSP orbit region. The median values from the models, obtained from Helios and OMNI measurements, are extrapolated to the PSP region for SSN values between solar minimum and maximum, that is, $0 \le ssn \le 200$. The lower edges of the shaded areas correspond to solar minimum, the upper edges to solar maximum. Also plotted are the radial dependencies of the analytic DQCS magnetic field model for solar minimum from \citet{Banaszkiewicz1998}, the slow wind velocity models from \citet{Sheeley1997} and \citet{Wang2000}, the density model from \citet{Leblanc1998} and the range of temperature measurements from \citet{Billings1959} and \citet{Liebenberg1975}.}

@@

        For determining solar-activity- and solar-distance-dependent relations\LEt{Please check that I have retained your intended meaning.} for the median and mean solar-wind values, we could have used the simpler approach of combining the radial dependence of averaged Helios data with averaged \SI{1}{\au} OMNI data scaled with the SSN. It is expected that the results of a simpler analysis would have similar distance scaling results, as can be inferred from the exponents in Tables~\ref{tab:mean_median_fit_parameter} and \ref{tab:extrapolation_model_fit_parameters}. However, in our study we are not only interested in averages but rather in bulk distributions, that is, the whole range of values that might occur. For the determination of the frequency distributions the use of the more complex fit model is important, because the distance between median and mean values determines the width of the lognormal distributions.
